Manuscript Draft

### 2019-01-09 manuscript.Rmd

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

@@

@@

```{r ccldb, echo=FALSE, include=FALSE}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
```

@@

```{r grid-dg-wilcox, echo = FALSE}
ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
```

Using cell-line-specific point mutation annotations from the CCLE, we compared z-score-transformed area under the dose-response curve (AUC) distributions between mutant and wildtype cell lines for all G2P genes for each G2P drug tested in the CCLE, CTRP, and GDSC drug screens (*Figures 1B and 2A*). At uncorrected $\alpha = 0.05$, Wilcoxon rank-sum tests revealed that in CCLE, the AUC distributions of mutant and wildtype groups for `r nrow(filter(ccle_signif_g2p_grid, p < 0.05))` out of `r nrow(ccle_signif_g2p_grid)` (`r printnum(nrow(filter(ccle_signif_g2p_grid, p < 0.05)) / nrow(ccle_signif_g2p_grid) * 100)`%) tested drug-gene combinations were significantly different, `r nrow(filter(ctrp_signif_g2p_grid, p < 0.05))` out of `r nrow(ctrp_signif_g2p_grid)` (`r printnum(nrow(filter(ctrp_signif_g2p_grid, p < 0.05)) / nrow(ctrp_signif_g2p_grid) * 100)`%) were significantly different for CTRP, and `r nrow(filter(gdsc_signif_g2p_grid, p < 0.05))` out of `r nrow(gdsc_signif_g2p_grid)` (`r printnum(nrow(filter(gdsc_signif_g2p_grid, p < 0.05)) / nrow(gdsc_signif_g2p_grid) * 100)`%) for GDSC. We observe little concordance between the number of interpretations for each G2P association and the significance of mutation status on AUC for said G2P association, suggesting that mutation status alone is not a robust predictor of clinically actionable genetic targets. This led us to examine the correlations between AUC and gene expression and AUC and copy number, two other genomic features that are known to associate with cancer and drug response.

@@

In all three datasets for both level-A G2P associations and null distribution associations, there are more significant |r| values when using gene expression as the genomic feature than when using copy number (Table 2), which corresponds with previous work showing that gene expression is an informative molecular feature when assessing drug sensitivity [@jang2014; @sirota2011]. Additionally, in all cases, there are more significant r values when using copy number than there are significant p-values from the mutation status comparisons. Though observational, these results suggest an order of predictive robustness for these three genomic features when used in isolation: Gene expression > copy number > mutation status. Additionally, across all three genomic features, CTRP consistently identified a higher percentage of its possible level-A G2P associations than GDSC, and similarly, GDSC identified a higher percentage than CCLE. In the CCLE-GDSC debate, inconsistency between the datasets was attributed to the groups using two very different assays for generating AUC measurements though it was undecided which was more accurate[@haibe-kains2013; @jang2014; @weinstein2013]. Our results provide very preliminary evidence that GDSC is more accurate than CCLE in regards to reproducing clinical associations. However, CTRP may be even more accurate than both GDSC and CCLE.

@@

```{r grid-dgl-wilcox, echo = FALSE}
ccle_signif_g2p_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
ccle_signif_grid_lineage <- readRDS("./data_munging/rds/ccle_signif_grid_lineage.rds")

ctrp_signif_g2p_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
ctrp_signif_grid_lineage <- readRDS("./data_munging/rds/ctrp_signif_grid_lineage.rds")

gdsc_signif_g2p_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
gdsc_signif_grid_lineage <- readRDS("./data_munging/rds/gdsc_signif_grid_lineage.rds")
```

@@

**BAD ANALYSIS, keep for reference:** These significant associations did not hold after Bonferroni correction in CTRP ($n = 48$, $\alpha_{adjusted} = 0.001$) or GDSC ($n = 39$, $\alpha_{adjusted} = 0.001$). When using gene expression as the genomic feature, three CCLE DGLs had significant correlations at an adjusted $\alpha =  0.006$ ($n = 9$). Two of these three DGLs were also significantly correlated with copy number ($n = 9$, $\alpha_{adjusted} = 0.006$). In CTRP, AUC of 19 DGLs significantly correlated with gene expression, 12 of which were significant after correction ($n = 61$, $\alpha_{adjusted} = 0.0008$). Copy number correlations were significant for 12 DGLs, with three DGLs significant post-correction ($n = 61$, $\alpha_{adjusted} = 0.0008$). AUC of 16 DGLs in GDSC significantly correlated with gene expression. At an adjusted $\alpha$ of 0.0009 ($n = 53$) eight of them were significant. Fewer DGLs were significant when correlating AUC with copy number with 6 significant at an $\alpha$ of 0.05 and 1 significant at an adjusted $\alpha$ of 0.0009 ($n = 53$).

@@

The CRISPR screen from the Broad Institute reports gene dependency using CERES scores [@meyers2017], which are generated from sgRNA depletion scores and eliminate bias arising from the effect of copy number variation on Cas9 DNA cleavage. The lower the CERES score, the higher the likelihood that the gene is essential in the associated cell line. Scores are scaled per cell line such that a score of 0 is the median effect of nonessential genes and -1 is the median effect of common core essential genes. Of the 34 G2P genes, only G6PD was not targeted by an sgRNA in the CRISPR dataset.

When we don’t account for lineage, Wilcoxon rank-sum tests revealed that mutations in eight of the 33 (24.2%) G2P genes significantly correlated with CERES score ($\alpha = 0.05$): KRAS ($p < 0.001$), BRAF ($p < 0.001$), BRCA2  ($p < 0.001$), CYP19A1 ($p = 0.004$), TSC2 ($p = 0.005$), JAK2 ($p = 0.031$), DPYD ($p = 0.042$), and UGT1A1 ($p = 0.048$). Gene expression (RPKM) of 17 (51.5%) G2P genes significantly correlated with CERES score, as did copy number (log2-ratio) for 14 (42.4%) G2P genes (Figure ##2). For the purpose of comparison to the drug screen data, if we treat CRISPR knockdown like a small-molecule drug, then these results are similar to the results from CCLE, CTRP, and GDSC (Table 2).

@@

Wilcoxon tests were computed using the ggpubr (`r ggpubr_v`) R package [@kassambara2018].

For computing Pearson correlation coefficients, if there were three or fewer data points, then the correlation was not done and a value of NA was supplied.

>>>>

floatsintext      : no
figuresintext     : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

```{r functions, echo=FALSE, include=FALSE}
e2exp_converter <- function(string) {
  e <- formatC(as.numeric(string), format = "e", digits = 0)
  exp <- sub("(\\d)e-0(\\d)", "$p_k < \\1^{-\\2}$", e, perl = TRUE)
  return(exp)
}
```

@@

# Wilcoxon test results
ccle_signif_g2p <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_g2p_exp <- e2exp_converter(ccle_signif_g2p$FDR_cutoff[1])
ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
ccle_signif_g2p_grid_exp <- e2exp_converter(ccle_signif_g2p_grid$FDR_cutoff[1])
ccle_signif_g2p_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
ccle_signif_g2p_lineage_exp <- e2exp_converter(ccle_signif_g2p_lineage$FDR_cutoff[1])
ccle_signif_grid_lineage <- readRDS("./data_munging/rds/ccle_signif_grid_lineage.rds")
ccle_signif_grid_lineage_exp <- e2exp_converter(ccle_signif_grid_lineage$FDR_cutoff[1])

ctrp_signif_g2p <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_g2p_exp <- e2exp_converter(ctrp_signif_g2p$FDR_cutoff[1])
ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
ctrp_signif_g2p_grid_exp <- e2exp_converter(ctrp_signif_g2p_grid$FDR_cutoff[1])
ctrp_signif_g2p_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
ctrp_signif_g2p_lineage_exp <- e2exp_converter(ctrp_signif_g2p_lineage$FDR_cutoff[1])
ctrp_signif_grid_lineage <- readRDS("./data_munging/rds/ctrp_signif_grid_lineage.rds")
ctrp_signif_grid_lineage_exp <- e2exp_converter(ctrp_signif_grid_lineage$FDR_cutoff[1])

gdsc_signif_g2p <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_g2p_exp <- e2exp_converter(gdsc_signif_g2p$FDR_cutoff[1])
gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
gdsc_signif_g2p_grid_exp <- e2exp_converter(gdsc_signif_g2p_grid$FDR_cutoff[1])
gdsc_signif_g2p_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
gdsc_signif_g2p_lineage_exp <- e2exp_converter(gdsc_signif_g2p_lineage$FDR_cutoff[1])
gdsc_signif_grid_lineage <- readRDS("./data_munging/rds/gdsc_signif_grid_lineage.rds")
gdsc_signif_grid_lineage_exp <- e2exp_converter(gdsc_signif_grid_lineage$FDR_cutoff[1])

# CCLE: GE/CN correlations
ccle_cor_res <- ccle_data_g2p_grid %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$p.value)
ccle_cor_res <- merge(g2p_druggene_tally, ccle_cor_res, by = "Drug_Gene", all.y = TRUE)
ccle_cor_res$n <- ifelse(is.na(ccle_cor_res$n), 0, ccle_cor_res$n)
ccle_cor_res$InG2P <- ifelse(ccle_cor_res$n != 0, "Yes", "No")
ccle_cor_res$Dataset <- "CCLE"

# CTRP: GE/CN correlations
ctrp_cor_res <- ctrp_data_g2p_grid %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$p.value)
ctrp_cor_res <- merge(g2p_druggene_tally, ctrp_cor_res, by = "Drug_Gene", all.y = TRUE)
ctrp_cor_res$n <- ifelse(is.na(ctrp_cor_res$n), 0, ctrp_cor_res$n)
ctrp_cor_res$InG2P <- ifelse(ctrp_cor_res$n != 0, "Yes", "No")
ctrp_cor_res$Dataset <- "CTRP"

# GDSC: GE/CN correlations
gdsc_cor_res <- gdsc_data_g2p_grid %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$p.value)
gdsc_cor_res <- merge(g2p_druggene_tally, gdsc_cor_res, by = "Drug_Gene", all.y = TRUE)
gdsc_cor_res$n <- ifelse(is.na(gdsc_cor_res$n), 0, gdsc_cor_res$n)
gdsc_cor_res$InG2P <- ifelse(gdsc_cor_res$n != 0, "Yes", "No")
gdsc_cor_res$Dataset <- "GDSC"

@@

@@

Using cell-line-specific point mutation annotations from the CCLE, we compared z-score-transformed area under the dose-response curve (AUC) distributions between mutant and wildtype cell lines for all G2P genes for each G2P drug tested in the CCLE, CTRP, and GDSC drug screens (*Figures 1B and 2A*). At an unadjusted $\alpha = 0.05$, Wilcoxon rank-sum tests revealed that in CCLE, the AUC distributions of mutant and wildtype groups for `r nrow(filter(ccle_signif_g2p_grid, p < 0.05))` out of `r nrow(ccle_signif_g2p_grid)` (`r printnum(nrow(filter(ccle_signif_g2p_grid, p < 0.05)) / nrow(ccle_signif_g2p_grid) * 100)`%) tested drug-gene combinations were significantly different, `r nrow(filter(ctrp_signif_g2p_grid, p < 0.05))` out of `r nrow(ctrp_signif_g2p_grid)` (`r printnum(nrow(filter(ctrp_signif_g2p_grid, p < 0.05)) / nrow(ctrp_signif_g2p_grid) * 100)`%) were significantly different for CTRP, and `r nrow(filter(gdsc_signif_g2p_grid, p < 0.05))` out of `r nrow(gdsc_signif_g2p_grid)` (`r printnum(nrow(filter(gdsc_signif_g2p_grid, p < 0.05)) / nrow(gdsc_signif_g2p_grid) * 100)`%) for GDSC.

Controlling for a false discovery rate (FDR) < 5%, selumetinib-*BRAF* ($p$ `r printp(ccle_signif_g2p_grid[ccle_signif_g2p_grid$Drug_Gene == "selumetinib_BRAF",]$p.adj)`) was the only significant drug-gene combination tested in CCLE (`r ccle_signif_g2p_grid_exp`). CTRP (`r ctrp_signif_g2p_grid_exp`) yielded `r nrow(ctrp_signif_g2p_grid)` significant combinations, one of which was selumetinib-*BRAF*.

We observe little concordance between the number of interpretations for each G2P association and the significance of mutation status on AUC for said G2P association, suggesting that mutation status alone is not a robust predictor of clinically actionable genetic targets. This led us to examine the correlations between AUC and gene expression and AUC and copy number, two other genomic features that are known to associate with cancer and drug response.

We computed Spearman correlation coefficients ($r_s$) between AUC of each drug-gene association and gene expression (RPKM) and copy number (log$_2$ ratio) of the gene in the association

@@

In all three datasets for both level-A G2P associations and null distribution associations, there are more significant |r| values when using gene expression as the genomic feature than when using copy number (Table 2), which corresponds with previous work showing that gene expression is an informative molecular feature when assessing drug sensitivity [@jang2014; @sirota2011]. Additionally, in all cases, there are more significant r values when using copy number than there are significant p-values from the mutation status comparisons. Though observational, these results suggest an order of predictive robustness for these three genomic features when used in isolation: Gene expression > copy number > mutation status. Additionally, across all three genomic features, CTRP consistently identified a higher percentage of its possible level-A G2P associations than GDSC, and similarly, GDSC identified a higher percentage than CCLE. In the CCLE-GDSC debate, inconsistency between the datasets was attributed to the groups using two very different assays for generating AUC measurements though it was undecided which was more accurate [@haibe-kains2013; @jang2014; @weinstein2013]. Our results provide very preliminary evidence that GDSC is more accurate than CCLE in regards to reproducing clinical associations. However, CTRP may be even more accurate than both GDSC and CCLE.

@@

@@

@@

```{r crispr-data, echo = FALSE, warning = FALSE, message = FALSE}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Hugo_Symbol %in% g2p_genes)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p$Mutation_Status_Nonsilent <- factor(crispr_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Wilcoxon test results
crispr_signif_g2p_gene <- readRDS("./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_g2p_gene_exp <- e2exp_converter(crispr_signif_g2p_gene$FDR_cutoff[1])
crispr_signif_g2p_gene_sig <- filter(crispr_signif_g2p_gene, !is.na(p.signif.adj))

# GE/CN correlations
crispr_spearman_res_GE <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>% summarize(p = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$p.value, corr = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Metric = "Gene Expression")
crispr_spearman_res_GE_sig <- filter(crispr_spearman_res_GE, p < 0.05)

crispr_spearman_res_CN <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$p.value, corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Metric = "Copy Number")
crispr_spearman_res_CN_sig <- filter(crispr_spearman_res_CN, p < 0.05)
```

The CRISPR screen from the Broad Institute reports gene dependency using CERES scores [@meyers2017], which are generated from sgRNA depletion scores and eliminate bias arising from the effect of copy number variation on Cas9 DNA cleavage. The lower the CERES score, the higher the likelihood that the gene is essential in the associated cell line. Scores are scaled per cell line such that a score of 0 is the median effect of nonessential genes and -1 is the median effect of common core essential genes. Of the `r length(g2p_genes)` G2P genes, only *`r setdiff(g2p_genes, as.character(crispr_signif_g2p_gene$Hugo_Symbol))`* was not targeted by an sgRNA in the CRISPR dataset.

When we don’t account for lineage, Wilcoxon rank-sum tests (`r crispr_signif_g2p_gene_exp`, FDR < 5%) revealed that mutations in `r nrow(crispr_signif_g2p_gene_sig)` genes significantly correlated with CERES score: *`r as.character(crispr_signif_g2p_gene_sig$Hugo_Symbol)[1]`* ($p$ `r printp(crispr_signif_g2p_gene_sig$p.adj[1])`) and *`r as.character(crispr_signif_g2p_gene_sig$Hugo_Symbol)[2]`* ($p$ `r printp(crispr_signif_g2p_gene_sig$p.adj[2])`).

Gene expression (RPKM) of `r nrow(crispr_spearman_res_GE_sig)` out of `r nrow(crispr_spearman_res_GE)` (`r printnum(nrow(crispr_spearman_res_GE_sig) / nrow(crispr_spearman_res_GE) * 100)`%) G2P genes significantly correlated with CERES score, as did copy number (log2-ratio) for `r nrow(crispr_spearman_res_CN_sig)` out of `r nrow(crispr_spearman_res_CN)` (`r printnum(nrow(crispr_spearman_res_CN_sig) / nrow(crispr_spearman_res_CN) * 100)`%) G2P genes.

For the purpose of comparison to the drug screen data, if we treat CRISPR knockdown like a small-molecule drug, then these results are similar to the results from CCLE, CTRP, and GDSC (Table 2).

@@

Wilcoxon tests were conducted using the `compare_means` function in the ggpubr (`r ggpubr_v`) R package [@kassambara2018]. `compare_means` does Benjamini-Hochberg p-value adjustment using the `p,adjust` function from the `stats` base R package.

For computing Spearman correlation coefficients. If there were 3 or fewer data points, then the correlation was not done and a value of NA was supplied.

### 2019-01-09 manuscript.tex

Using cell-line-specific point mutation annotations from the CCLE, we
compared z-score-transformed area under the dose-response curve (AUC)
distributions between mutant and wildtype cell lines for all G2P genes
for each G2P drug tested in the CCLE, CTRP, and GDSC drug screens
(\emph{Figures 1B and 2A}). At uncorrected \(\alpha = 0.05\), Wilcoxon
rank-sum tests revealed that in CCLE, the AUC distributions of mutant
and wildtype groups for 18 out of 170 (10.59\%) tested drug-gene
combinations were significantly different, 135 out of 918 (14.71\%) were
significantly different for CTRP, and 89 out of 850 (10.47\%) for GDSC.
We observe little concordance between the number of interpretations for
each G2P association and the significance of mutation status on AUC for
said G2P association, suggesting that mutation status alone is not a
robust predictor of clinically actionable genetic targets. This led us
to examine the correlations between AUC and gene expression and AUC and
copy number, two other genomic features that are known to associate with
cancer and drug response.

@@

In all three datasets for both level-A G2P associations and null
distribution associations, there are more significant
\textbar{}r\textbar{} values when using gene expression as the genomic
feature than when using copy number (Table 2), which corresponds with
previous work showing that gene expression is an informative molecular
feature when assessing drug sensitivity (Jang et al., 2014; Sirota et
al., 2011). Additionally, in all cases, there are more significant r
values when using copy number than there are significant p-values from
the mutation status comparisons. Though observational, these results
suggest an order of predictive robustness for these three genomic
features when used in isolation: Gene expression \textgreater{} copy
number \textgreater{} mutation status. Additionally, across all three
genomic features, CTRP consistently identified a higher percentage of
its possible level-A G2P associations than GDSC, and similarly, GDSC
identified a higher percentage than CCLE. In the CCLE-GDSC debate,
inconsistency between the datasets was attributed to the groups using
two very different assays for generating AUC measurements though it was
undecided which was more accurate(Haibe-Kains et al., 2013; Jang et al.,
2014; Weinstein \& Lorenzi, 2013). Our results provide very preliminary
evidence that GDSC is more accurate than CCLE in regards to reproducing
clinical associations. However, CTRP may be even more accurate than both
GDSC and CCLE.

@@

Prior to correction for an FDR of 5\%, no DGLs were identified in the
CCLE screen when using mutation status as a correlate (\(n =\) 7, 2 DGLs
were not analyzed due to lack of mutation annotations in \emph{UGT1A1}
in \enquote{cancer} lineages), but CTRP (\(n =\) 48) and GDSC (\(n =\)
39) identified 6 and 5, respectively, in skin cancer, lung cancer, and
cancer.

\textbf{BAD ANALYSIS, keep for reference:} These significant
associations did not hold after Bonferroni correction in CTRP
(\(n = 48\), \(\alpha_{adjusted} = 0.001\)) or GDSC (\(n = 39\),
\(\alpha_{adjusted} = 0.001\)). When using gene expression as the
genomic feature, three CCLE DGLs had significant correlations at an
adjusted \(\alpha = 0.006\) (\(n = 9\)). Two of these three DGLs were
also significantly correlated with copy number (\(n = 9\),
\(\alpha_{adjusted} = 0.006\)). In CTRP, AUC of 19 DGLs significantly
correlated with gene expression, 12 of which were significant after
correction (\(n = 61\), \(\alpha_{adjusted} = 0.0008\)). Copy number
correlations were significant for 12 DGLs, with three DGLs significant
post-correction (\(n = 61\), \(\alpha_{adjusted} = 0.0008\)). AUC of 16
DGLs in GDSC significantly correlated with gene expression. At an
adjusted \(\alpha\) of 0.0009 (\(n = 53\)) eight of them were
significant. Fewer DGLs were significant when correlating AUC with copy
number with 6 significant at an \(\alpha\) of 0.05 and 1 significant at
an adjusted \(\alpha\) of 0.0009 (\(n = 53\)).

@@

The CRISPR screen from the Broad Institute reports gene dependency using
CERES scores (Meyers et al., 2017), which are generated from sgRNA
depletion scores and eliminate bias arising from the effect of copy
number variation on Cas9 DNA cleavage. The lower the CERES score, the
higher the likelihood that the gene is essential in the associated cell
line. Scores are scaled per cell line such that a score of 0 is the
median effect of nonessential genes and -1 is the median effect of
common core essential genes. Of the 34 G2P genes, only G6PD was not
targeted by an sgRNA in the CRISPR dataset.

When we don't account for lineage, Wilcoxon rank-sum tests revealed that
mutations in eight of the 33 (24.2\%) G2P genes significantly correlated
with CERES score (\(\alpha = 0.05\)): KRAS (\(p < 0.001\)), BRAF
(\(p < 0.001\)), BRCA2 (\(p < 0.001\)), CYP19A1 (\(p = 0.004\)), TSC2
(\(p = 0.005\)), JAK2 (\(p = 0.031\)), DPYD (\(p = 0.042\)), and UGT1A1
(\(p = 0.048\)). Gene expression (RPKM) of 17 (51.5\%) G2P genes
significantly correlated with CERES score, as did copy number
(log2-ratio) for 14 (42.4\%) G2P genes (Figure \#\#2). For the purpose
of comparison to the drug screen data, if we treat CRISPR knockdown like
a small-molecule drug, then these results are similar to the results
from CCLE, CTRP, and GDSC (Table 2).

@@

Wilcoxon tests were computed using the ggpubr (v0.1.7.999) R package
(Kassambara, 2018).

For computing Pearson correlation coefficients, if there were three or
fewer data points, then the correlation was not done and a value of NA
was supplied.

>>>>

Using cell-line-specific point mutation annotations from the CCLE, we
compared z-score-transformed area under the dose-response curve (AUC)
distributions between mutant and wildtype cell lines for all G2P genes
for each G2P drug tested in the CCLE, CTRP, and GDSC drug screens
(\emph{Figures 1B and 2A}). At an unadjusted \(\alpha = 0.05\), Wilcoxon
rank-sum tests revealed that in CCLE, the AUC distributions of mutant
and wildtype groups for 18 out of 170 (10.59\%) tested drug-gene
combinations were significantly different, 135 out of 918 (14.71\%) were
significantly different for CTRP, and 89 out of 850 (10.47\%) for GDSC.

Controlling for a false discovery rate (FDR) \textless{} 5\%,
selumetinib-\emph{BRAF} (\(p\) \textless{} .001) was the only
significant drug-gene combination tested in CCLE (\(p_k < 8^{-4}\)).
CTRP (\(p_k < 7^{-4}\)) yielded 918 significant combinations, one of
which was selumetinib-\emph{BRAF}.

We observe little concordance between the number of interpretations for
each G2P association and the significance of mutation status on AUC for
said G2P association, suggesting that mutation status alone is not a
robust predictor of clinically actionable genetic targets. This led us
to examine the correlations between AUC and gene expression and AUC and
copy number, two other genomic features that are known to associate with
cancer and drug response.

We computed Spearman correlation coefficients (\(r_s\)) between AUC of
each drug-gene association and gene expression (RPKM) and copy number
(log\(_2\) ratio) of the gene in the association

@@

In all three datasets for both level-A G2P associations and null
distribution associations, there are more significant
\textbar{}r\textbar{} values when using gene expression as the genomic
feature than when using copy number (Table 2), which corresponds with
previous work showing that gene expression is an informative molecular
feature when assessing drug sensitivity (Jang et al., 2014; Sirota et
al., 2011). Additionally, in all cases, there are more significant r
values when using copy number than there are significant p-values from
the mutation status comparisons. Though observational, these results
suggest an order of predictive robustness for these three genomic
features when used in isolation: Gene expression \textgreater{} copy
number \textgreater{} mutation status. Additionally, across all three
genomic features, CTRP consistently identified a higher percentage of
its possible level-A G2P associations than GDSC, and similarly, GDSC
identified a higher percentage than CCLE. In the CCLE-GDSC debate,
inconsistency between the datasets was attributed to the groups using
two very different assays for generating AUC measurements though it was
undecided which was more accurate (Haibe-Kains et al., 2013; Jang et
al., 2014; Weinstein \& Lorenzi, 2013). Our results provide very
preliminary evidence that GDSC is more accurate than CCLE in regards to
reproducing clinical associations. However, CTRP may be even more
accurate than both GDSC and CCLE.

@@

Prior to correction for an FDR of 5\%, no DGLs were identified in the
CCLE screen when using mutation status as a correlate (\(n =\) 7, 2 DGLs
were not analyzed due to lack of mutation annotations in \emph{UGT1A1}
in \enquote{cancer} lineages), but CTRP (\(n =\) 48) and GDSC (\(n =\)
39) identified 6 and 5, respectively, in skin cancer, lung cancer, and
cancer. After Benjamini-Hochberg FDR correction, none of the datasets
yielded significant DGL interactions.

@@

The CRISPR screen from the Broad Institute reports gene dependency using
CERES scores (Meyers et al., 2017), which are generated from sgRNA
depletion scores and eliminate bias arising from the effect of copy
number variation on Cas9 DNA cleavage. The lower the CERES score, the
higher the likelihood that the gene is essential in the associated cell
line. Scores are scaled per cell line such that a score of 0 is the
median effect of nonessential genes and -1 is the median effect of
common core essential genes. Of the 34 G2P genes, only \emph{G6PD} was
not targeted by an sgRNA in the CRISPR dataset.

When we don't account for lineage, Wilcoxon rank-sum tests
(\(p_k < 5^{-3}\), FDR \textless{} 5\%) revealed that mutations in 2
genes significantly correlated with CERES score: \emph{KRAS} (\(p\)
\textless{} .001) and \emph{BRAF} (\(p\) \textless{} .001).

Gene expression (RPKM) of 12 out of 33 (36.36\%) G2P genes significantly
correlated with CERES score, as did copy number (log2-ratio) for 12 out
of 33 (36.36\%) G2P genes.

For the purpose of comparison to the drug screen data, if we treat
CRISPR knockdown like a small-molecule drug, then these results are
similar to the results from CCLE, CTRP, and GDSC (Table 2).

@@

Wilcoxon tests were conducted using the \texttt{compare\_means} function
in the ggpubr (v0.1.7.999) R package (Kassambara, 2018).
\texttt{compare\_means} does Benjamini-Hochberg p-value adjustment using
the \texttt{p,adjust} function from the \texttt{stats} base R package.

For computing Spearman correlation coefficients. If there were 3 or
fewer data points, then the correlation was not done and a value of NA
was supplied.
